---
title: 'Capstone Project 1: Shark Attacks in Coastal Waters.'
author: "Randy Baicich"
date: "06/16/2023"
output: html_document
---

Starting Hypothesis:

Based on initial observations and assumptions, the following hypotheses are proposed: A select few shark species will be responsible for the majority of shark attacks and fatalities, rather than a wide variety of species. The majority of shark attacks will occur at specific times, indicating a temporal pattern in shark-human interactions. The objective is to test these hypotheses and gain deeper insights into the key factors influencing shark attacks and fatalities.

## Choose and Install necessary packages and libraries.

```{r}
#Load required libraries
library(dplyr)
library(ggplot2)
```


```{r}
#Set the file path
file_path <- "C:/Users/RedneckRandy/Downloads/archive (15)/GSAF5.csv"

#Import the CSV file
data <- read.csv(file_path, stringsAsFactors = FALSE)
```


```{r}
#Clean the CSV file
cleaned_data <- data %>%
  select(-matches("^(u|U)nnamed")) %>%
  na.omit()
print(cleaned_data)
```


```{r}
#Calculate the number of occurrences for each species
species_occurrences <- cleaned_data %>%
  group_by(`Species`) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))
print(species_occurrences)
```


```{r}
#Convert "Time" column to numeric
cleaned_data$Time <- as.numeric(as.character(cleaned_data$Time))

#Calculate the average of "Time"
average_time <- cleaned_data %>%
  summarize(Average_Time = mean(Time, na.rm = TRUE))

#Print the average time
cat("Average Time: ", average_time$Average_Time, "\n")

```
```{r}
#Get the top 10 locations based on occurrence count
top_locations <- cleaned_data %>%
  group_by(Location) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count)) %>%
  top_n(10)
print(top_locations)
```

```{r}
#Calculate the number of occurrences for each country
country_occurrences <- cleaned_data %>%
  group_by(Country) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))
print(country_occurrences)
```
```{r}
#Calculate the number of occurrences for each activity
activity_occurrences <- cleaned_data %>%
  group_by(Activity) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))
print(activity_occurrences)
```


#Visualize the data


```{r}
#Graph 1: Top 10 Countries vs Number of Occurrences
top_10_countries <- head(country_occurrences, 10)  #Select the top 10 countries

graph3 <- ggplot(top_10_countries, aes(x = Country, y = Count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(x = "Country", y = "Number of Occurrences") +
  ggtitle("Number of Occurrences by Country (Top 10)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(graph1)
```

```{r}
#Print the analysis results
print(average_time)

print(species_occurrences)

print(country_occurrences)

print(activity_occurrences)

print(top_locations)

```


## Summary/Conclusion.


The analysis of shark attack data reveals important patterns and considerations. The average time of attack is approximately 1050, highlighting a specific timeframe when shark-human interactions are more likely to occur. The White shark stands out as the species with the highest number of occurrences, although caution is needed when interpreting categories involving unconfirmed shark involvement. The United States has the highest number of reported shark attacks, emphasizing the need for safety measures in coastal waters. Surfing and swimming are identified as the most dangerous activities in terms of shark attacks, underscoring the importance of adhering to safety guidelines. Overall, these findings emphasize the significance of understanding species behavior, geographic patterns, and promoting safe practices to mitigate risks in shark-human interactions.
